Trajectories

### 2021-09-15 Rmd/report.Rmd
mode report to Markdown

---
title: "Trajectories"
output:
  html_document:
    df_print: paged
    toc: true
    toc_depth: 4
    toc_float: true
bibliography: "trajectories.bib"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  message = TRUE,
  warning = TRUE, 
  cache.extra = knitr::rand_seed,
  out.width = "100%",
  results = "asis",
  dev.args = list(png = list(type = "cairo"))
)
options(knitr.kable.NA = '-',
        ggplot2.discrete.fill = wesanderson::wes_palettes$Darjeeling1,
        ggplot2.discrete.colour = wesanderson::wes_palettes$Darjeeling1,
        ggplot2.continuous.fill = ggplot2::scale_color_gradient,
        ggplot2.continuous.colour = ggplot2::scale_color_gradient
)

@@

# load packages
library(targets)
library(tidyverse)
library(gt)
library(lubridate)
library(wesanderson)
library(brms)
library(htmltools)
library(tidybayes)
library(scales)
library(patchwork)
library(janitor)
library(here)

# set params
theme_set(theme_custom())

@@

```{r participants_lp, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
labs <- c("0-10%", "10-20%", "20-30%", "30-40%", "40-50%")
participants %>% 
  mutate(
    age = round(age),
    doe_2 = ifelse(dominant_language=="Catalan", doe_spanish, doe_catalan),
    doe_2 = cut(doe_2, breaks = seq(0, 0.5, 0.1), include.lowest = TRUE, labels = labs)
  ) %>% 
  count(age, doe_2) %>% 
  arrange(doe_2) %>% 
  mutate(doe_2 = fct_inorder(doe_2)) %>% 
  pivot_wider(names_from = doe_2, values_from = n, values_fill = 0) %>% 
  gt() %>% 
  tab_spanner(label = "Number of participants by DOE", columns = 2:6) %>% 
  data_color(
    columns = 2:6,
    colors = scales::col_numeric(
      palette = c("white", "orange"),
      domain = c(0, 11)
    )
  ) %>% 
  cols_label(
    age = md("**Age<br>(months)**")
  )

@@

vocabulary %>% 
  filter(type=="Understands", modality != "Te") %>% 
  ggplot(aes(age, prop, colour = lp, fill = lp)) +
  facet_wrap(~modality) +
  geom_point(size = 0.5, alpha = 0.21) +
  geom_smooth(method = "glm", method.args = list(family = "binomial"), se = FALSE) +
  labs(x = "Age (months)", y = "Vocabulary size (%)", colour = "LP", fill = "LP") +
  scale_y_continuous(limits = c(0, 1), labels = scales::percent) +

  vocabulary %>% 
  filter(type=="Understands", modality != "Te") %>% 
  ggplot(aes(age, count, colour = lp, fill = lp)) +
  facet_wrap(~modality) +
  geom_point(size = 0.5, alpha = 0.21) +
  geom_smooth(method = "glm", method.args = list(family = "poisson"), se = FALSE) +
  labs(x = "Age (months)", y = "Vocabulary size (words)", colour = "LP", fill = "LP") +

@@

vocabulary %>% 
  filter(type=="Produces", modality != "Te") %>% 
  ggplot(aes(age, prop, colour = lp, fill = lp)) +
  facet_wrap(~modality) +
  geom_point(size = 0.5, alpha = 0.21) +
  geom_smooth(method = "glm", method.args = list(family = "binomial"), se = FALSE) +
  labs(x = "Age (months)", y = "Vocabulary size (%)", colour = "LP", fill = "LP") +
  scale_y_continuous(limits = c(0, 1), labels = scales::percent) +

  vocabulary %>% 
  filter(type=="Understands", modality != "Te") %>% 
  ggplot(aes(age, count, colour = lp, fill = lp)) +
  facet_wrap(~modality) +
  geom_point(size = 0.5, alpha = 0.21) +
  geom_smooth(method = "glm", method.args = list(family = "poisson"), se = FALSE) +
  labs(x = "Age (months)", y = "Vocabulary size (words)", colour = "LP", fill = "LP") +

@@

```{r vocabulary_long_comprehension, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
vocabulary_long %>% 
  filter(type=="understands") %>% 
  ggplot(aes(value, reorder(id, age), colour = version, fill = version, shape = version)) +
  facet_grid(~scale) +
  geom_point() +
  labs(x= "Vocabulary size", y = "Participant (ordered by age, min-to-max)",
       colour = "Version", fill = "Version", shape = "Version") +
  scale_x_continuous(labels = scales::percent) +
  theme_bw() +
  theme(
    legend.position = "top",
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank()
  )
```

@@

```{r vocabulary_long_production, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
vocabulary_long %>% 
  filter(type=="produces") %>% 
  ggplot(aes(value, reorder(id, age), colour = version, fill = version, shape = version)) +
  facet_grid(~scale) +
  geom_point() +
  labs(x= "Vocabulary size", y = "Participant (ordered by age, min-to-max)",
       colour = "Version", fill = "Version", shape = "Version") +
  scale_x_continuous(labels = scales::percent) +
  theme_bw() +
  theme(
    legend.position = "top",
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank()
  )
```

@@

# emmeans
nd <- expand.grid(
  age = seq(min(responses$age), max(responses$age), length.out = 50),
  lp = c("Monolingual", "Bilingual"),
  cognate = c("Cognate", "Non-Cognate"),
  frequency_center = c(-1, 0, 1)
)

@@

```{r fixed_effects_table, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
summary(fit)$fixed %>% 
  as.data.frame() %>%
  rownames_to_column("term") %>% 
  mutate(
    Estimate = ifelse(term=="Intercept", inv_logit_scaled(Estimate), Estimate/4),
    `l-95% CI` = ifelse(term=="Intercept", inv_logit_scaled(`l-95% CI`), `l-95% CI`/4),
    `u-95% CI` = ifelse(term=="Intercept", inv_logit_scaled(`u-95% CI`), `u-95% CI`/4),
    Est.Error = ifelse(term=="Intercept", inv_logit_scaled(Est.Error), Est.Error/4)
  ) %>% 
  gt() %>% 
  fmt_percent(2:5) %>% 
  fmt_number(6:6, decimals = 2) %>% 
  fmt_number(7:8, decimals = 0) %>% 
  cols_merge(vars("l-95% CI", "u-95% CI"), pattern = "[{1}, {2}]") %>% 
  cols_label(
    term = md("**Predictor**"),
    Estimate = md("**Mean**"),
    Est.Error = md("**SEM**"),
    "l-95% CI" = md("**95\\% CrI**"),
    Rhat = md("**Rhat**"),
    Bulk_ESS = md("**Bulk ESS**"),
    Tail_ESS = md("**Tail ESS**")
  ) %>% 
  tab_footnote(
    footnote = "Transformed using the inverse logit to get the average probability of correct response",
    locations = cells_body(columns = "term", rows = term=="Intercept")
  ) %>% 
  tab_footnote(
    footnote = "Transformed using the divide-by-four- rule to get the maximum change in probability of correct response, associated with a unit increase in this variable.",
    locations = cells_body(columns = "term", rows = term %in% c("age", "frequency", "doe", "cognate1", "doe:cognate1"))
  ) %>% 
  tab_footnote(
    footnote = "ESS: Effective sample size",
    locations = cells_column_labels(columns = c("Bulk_ESS", "Tail_ESS"))
  ) %>% 
  cols_align(
    align = c("center"),
    columns = 2:4
  )

@@

post <- gather_draws(fit, `b_.*`, regex = TRUE) %>% 
  mutate(.value = ifelse(.variable=="b_Intercept", inv_logit_scaled(.value), .value/4))

@@

re_id <- ranef(fit)$id[,,1] %>% 
  as.data.frame() %>% 
  rownames_to_column("id") %>% 
  as_tibble() %>% 
  mutate_if(is.numeric, function(x) inv_logit_scaled(x + fixef(fit)[1])) %>% 
  left_join(distinct(fit$data, id, lp))

@@

re_te <- ranef(fit)$te[,,1] %>% 
  as.data.frame() %>% 
  rownames_to_column("te") %>% 
  as_tibble() %>% 
  mutate_if(is.numeric, function(x) inv_logit_scaled(x + fixef(fit)[1])) %>% 
  mutate(te = as.numeric(te)) %>% 
  left_join(distinct(fit$data, te, lp))

@@

nd <- expand.grid(
  age = seq(min(responses$age), max(responses$age), length.out = 50),
  lp = c("Monolingual", "Bilingual"),
  cognate = c("Cognate", "Non-Cognate"),
  frequency_center = c(-1,  0, 1)
)

m <- add_fitted_draws(nd, fit, re_formula = NA, n = 25) %>% 
  mutate(frequency_center = as.factor(paste0(frequency_center, " SD")))

@@

gather_draws(fit, `b_.*`, `sd_te__.*`, `cor_te__.*`, regex = TRUE) %>% 
  mutate(.chain = paste("Chain ", .chain)) %>% 
  ggplot(aes(.iteration, .value, colour = .chain)) +
  facet_wrap(~.variable, scales = "free_y") +
  annotate(geom = "rect", colour = NA, xmin = 0, xmax = 500, ymin = -Inf, ymax = Inf, alpha = 0.25, colour = "grey") +
  geom_line(alpha = 0.5) +
  labs(x = "Iteration", y = "Sample value", colour = "Chain") +
  theme_custom() +
  theme(
    legend.position = "top",
    legend.title = element_blank(),
    panel.grid = element_blank()
  )

>>>>

---
title: "Trajectories"
output:
  github_document:
    pandoc_args: --webtex
bibliography: "trajectories.bib"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  message = TRUE,
  echo = TRUE,
  warning = TRUE,
  dev.args = list(png = list(type = "cairo")),
  include = FALSE,
  out.width = "100%",
  results = "asis",
  dpi = 300
)
options(
  knitr.kable.NA = "-",
  knitr.duplicate.label = "allow",
  ggplot2.discrete.fill = wesanderson::wes_palettes$Darjeeling1,
  ggplot2.discrete.colour = wesanderson::wes_palettes$Darjeeling1,
  ggplot2.continuous.fill = ggplot2::scale_color_gradient,
  ggplot2.continuous.colour = ggplot2::scale_color_gradient
)

@@

# load packages
library(targets)
library(dplyr)
library(tidyr)
library(stringr)
library(gt)
library(lubridate)
library(wesanderson)
library(brms)
library(htmltools)
library(tidybayes)
library(scales)
library(patchwork)
library(tibble)
library(janitor)
library(here)

# set params
tar_load_globals()
theme_set(theme_custom())

@@

```{r participants_lp, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
labs <- c("0-10%", "10-20%", "20-30%", "30-40%", "40-50%")
a <- participants %>% 
  mutate(
    age = round(age),
    doe_2 = ifelse(dominant_language=="Catalan", doe_spanish, doe_catalan),
    doe_2 = cut(doe_2, breaks = seq(0, 0.5, 0.1), include.lowest = TRUE, labels = labs)
  ) %>% 
  count(age, doe_2) %>% 
  arrange(doe_2) %>% 
  mutate(doe_2 = fct_inorder(doe_2)) %>% 
  pivot_wider(names_from = doe_2, values_from = n, values_fill = 0) %>% 
  gt() %>% 
  tab_spanner(label = "Number of participants by DOE", columns = 2:6) %>% 
  data_color(
    columns = 2:6,
    colors = scales::col_numeric(
      palette = c("white", "orange"),
      domain = c(0, 11)
    )
  ) %>% 
  cols_label(
    age = md("**Age<br>(months)**")
  )

@@

d <- vocabulary$wide %>% 
  filter(type=="Understands", modality != "Te")

ggplot(d, aes(age, prop, colour = lp, fill = lp)) +
  facet_wrap(~modality) +
  geom_point(size = 0.5, alpha = 0.21) +
  geom_smooth(method = "glm", method.args = list(family = "binomial"), se = FALSE) +
  labs(x = "Age (months)", y = "Vocabulary size (%)", colour = "LP", fill = "LP") +
  scale_y_continuous(limits = c(0, 1), labels = scales::percent) +

  ggplot(d, aes(age, count, colour = lp, fill = lp)) +
  facet_wrap(~modality) +
  geom_point(size = 0.5, alpha = 0.21) +
  geom_smooth(method = "glm", method.args = list(family = "poisson"), se = FALSE) +
  labs(x = "Age (months)", y = "Vocabulary size (words)", colour = "LP", fill = "LP") +

@@

vocabulary$wide %>% 
  filter(type=="Produces", modality != "Te") %>% 
  ggplot(aes(age, prop, colour = lp, fill = lp)) +
  facet_wrap(~modality) +
  geom_point(size = 0.5, alpha = 0.21) +
  geom_smooth(method = "glm", method.args = list(family = "binomial"), se = FALSE) +
  labs(x = "Age (months)", y = "Vocabulary size (%)", colour = "LP", fill = "LP") +
  scale_y_continuous(limits = c(0, 1), labels = scales::percent) +

  vocabulary$wide %>% 
  filter(type=="Produces", modality != "Te") %>% 
  ggplot(aes(age, count, colour = lp, fill = lp)) +
  facet_wrap(~modality) +
  geom_point(size = 0.5, alpha = 0.21) +
  geom_smooth(method = "glm", method.args = list(family = "poisson"), se = FALSE) +
  labs(x = "Age (months)", y = "Vocabulary size (words)", colour = "LP", fill = "LP") +

@@

```{r vocabulary_long_comprehension, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
vocabulary$long %>% 
  filter(type=="understands") %>% 
  ggplot(aes(value, reorder(id, age), colour = version, fill = version, shape = version)) +
  facet_grid(~scale) +
  geom_point() +
  labs(x= "Vocabulary size", y = "Participant (ordered by age, min-to-max)",
       colour = "Version", fill = "Version", shape = "Version") +
  scale_x_continuous(labels = scales::percent) +
  theme_bw() +
  theme(
    legend.position = "top",
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank()
  )
```

@@

```{r vocabulary_long_production, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
vocabulary$long %>% 
  filter(type=="produces") %>% 
  ggplot(aes(value, reorder(id, age), colour = version, fill = version, shape = version)) +
  facet_grid(~scale) +
  geom_point() +
  labs(x= "Vocabulary size", y = "Participant (ordered by age, min-to-max)",
       colour = "Version", fill = "Version", shape = "Version") +
  scale_x_continuous(labels = scales::percent) +
  theme_bw() +
  theme(
    legend.position = "top",
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank()
  )
```

@@

# emmeans
nd <- expand.grid(
  age = seq(min(responses$age), max(responses$age), length.out = 50),
  lp = unique(participants$lp),
  cognate = unique(items$cognate),
  frequency_center = c(-1, 0, 1)
)

@@

```{r fixed_effects_table, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
summary(fits$fit_2)$fixed %>% 
  as.data.frame() %>%
  rownames_to_column("term") %>% 
  mutate(
    Estimate = ifelse(term=="Intercept", inv_logit_scaled(Estimate), Estimate/4),
    `l-95% CI` = ifelse(term=="Intercept", inv_logit_scaled(`l-95% CI`), `l-95% CI`/4),
    `u-95% CI` = ifelse(term=="Intercept", inv_logit_scaled(`u-95% CI`), `u-95% CI`/4),
    Est.Error = ifelse(term=="Intercept", inv_logit_scaled(Est.Error), Est.Error/4)
  ) %>% 
  gt() %>% 
  fmt_percent(2:5) %>% 
  fmt_number(6:6, decimals = 2) %>% 
  fmt_number(7:8, decimals = 0) %>% 
  cols_merge(vars("l-95% CI", "u-95% CI"), pattern = "[{1}, {2}]") %>% 
  cols_label(
    term = md("**Predictor**"),
    Estimate = md("**Mean**"),
    Est.Error = md("**SEM**"),
    "l-95% CI" = md("**95\\% CrI**"),
    Rhat = md("**Rhat**"),
    Bulk_ESS = md("**Bulk ESS**"),
    Tail_ESS = md("**Tail ESS**")
  ) %>% 
  tab_footnote(
    footnote = "Transformed using the inverse logit to get the average probability of correct response",
    locations = cells_body(columns = "term", rows = term=="Intercept")
  ) %>% 
  tab_footnote(
    footnote = "Transformed using the divide-by-four- rule to get the maximum change in probability of correct response, associated with a unit increase in this variable.",
    locations = cells_body(columns = "term", rows = term %in% c("age", "frequency", "doe", "cognate1", "doe:cognate1"))
  ) %>% 
  tab_footnote(
    footnote = "ESS: Effective sample size",
    locations = cells_column_labels(columns = c("Bulk_ESS", "Tail_ESS"))
  ) %>% 
  cols_align(
    align = c("center"),
    columns = 2:4
  )

@@

post <- gather_draws(fits$fit_2, `b_.*`, regex = TRUE) %>% 
  mutate(.value = ifelse(.variable=="b_Intercept", inv_logit_scaled(.value), .value/4))

@@

re_id <- ranef(fits$fit_2)$id[,,1] %>% 
  as.data.frame() %>% 
  rownames_to_column("id") %>% 
  as_tibble() %>% 
  mutate_if(is.numeric, function(x) inv_logit_scaled(x + fixef(fits$fit_2)[1])) %>% 
  left_join(distinct(fits$fit_2$data, id, lp))

@@

re_te <- ranef(fits$fit_2)$te[,,1] %>% 
  as.data.frame() %>% 
  rownames_to_column("te") %>% 
  as_tibble() %>% 
  mutate_if(is.numeric, function(x) inv_logit_scaled(x + fixef(fits$fit_2)[1])) %>% 
  mutate(te = as.numeric(te)) %>% 
  left_join(distinct(fits$fit_2$data, te, lp))

@@

nd <- expand.grid(
  age = seq(min(responses$age), max(responses$age), length.out = 50),
  lp = unique(participants$lp),
  cognate =unique(items$cognate),
  frequency_center = c(-1,  0, 1)
)

m <- add_fitted_draws(nd, fits$fit_2, re_formula = NA, n = 25) %>% 
  mutate(frequency_center = as.factor(paste0(frequency_center, " SD")))

@@

gather_draws(fits$fit_2, `b_.*`, `sd_te__.*`, `cor_te__.*`, regex = TRUE) %>% 
  mutate(.chain = paste("Chain ", .chain)) %>% 
  ggplot(aes(.iteration, .value, colour = .chain)) +
  facet_wrap(~.variable, scales = "free_y") +
  annotate(geom = "rect", colour = NA, xmin = 0, xmax = 500, ymin = -Inf, ymax = Inf, alpha = 0.25, colour = "grey") +
  geom_line(alpha = 0.5) +
  labs(x = "Iteration", y = "Sample value", colour = "Chain") +
  theme_custom() +
  theme(
    legend.position = "top",
    legend.title = element_blank(),
    panel.grid = element_blank()
  )
